---
title: "ARE 213 Problem Set 1b"
author: "Nick Depsky, Will Gorman, Peter Worley"
date: "October 12, 2018"
output:
  pdf_document: default
  html_notebook: default
---

```{r, include = F}
rm(list = ls())
library(pacman)
p_load("foreign","dplyr","magrittr","knitr","ggplot2", "corrplot", "stargazer")
theme_plot <- theme(
  legend.position = "right",
  panel.background = element_rect(fill = NA),
  panel.border = element_rect(fill = NA, color = "grey75"),
  axis.ticks = element_line(color = "grey85"),
  panel.grid.major = element_line(color = "grey95", size = 0.2),
  panel.grid.minor = element_line(color = "grey95", size = 0.2),
  legend.key = element_blank(),
  legend.title = element_blank(),
  legend.spacing.x = unit(0.3, "cm"))
```

Import Data
```{r}
#setwd("~/Dropbox/Berkeley_tings/Fall 2018/ARE213/Problem Sets/PS1")
#setwd("C:\\Users\\will-\\Desktop\\are213")
setwd("C:\\Users\\Will\\Desktop\\are213\\PS1b")
dat <- read.dta("ps1.dta")
#fix missing data
dat_drop <- dat %>% filter(herpes != 8 & tobacco != 9 & cigar != 99 & cigar6 != 6 & 
                  alcohol != 9 & drink != 99 & drink5 != 5 & wgain != 99)
```

# 1a - Misspecification bias


# 1b - Higher order specifications


```{r 1b}


```

#1c - Using LASSO
```{r 1c}


```



#2 - Propensity score description

#2a - Create propensity score

```{r, echo = F}


```

#2b - propensity score estimation

```{r}


```



#2c - propensity score reweighting



#2d - kernel density estimator


```{r}

```


#2e - bandwidth adjustments


```{r}

```

#2f - benefits and drawbacks of propensity method


#2g - Present and discuss results

```{r}

```

#3 - Blocking non-parametric approach

```{r}

```

#4 - low birth weights

```{r}

```

#5 - Summarize results


