---
title: "ARE 213 Problem Set 2a"
author: "Nick Depsky, Will Gorman, Peter Worley"
date: "October 26, 2018"
output:
  pdf_document: default
  html_notebook: default
---

```{r, include = F}
rm(list = ls())
library(pacman)
p_load("foreign","dplyr","magrittr","knitr","ggplot2", "corrplot", "stargazer", "glmnet","splines")
theme_plot <- theme(
  legend.position = "right",
  panel.background = element_rect(fill = NA),
  panel.border = element_rect(fill = NA, color = "grey75"),
  axis.ticks = element_line(color = "grey85"),
  panel.grid.major = element_line(color = "grey95", size = 0.2),
  panel.grid.minor = element_line(color = "grey95", size = 0.2),
  legend.key = element_blank(),
  legend.title = element_blank(),
  legend.spacing.x = unit(0.3, "cm"))

```

# 1a - Show betas numerically identical

# 1b - Show standard errors numerically identical


# 2a - Show fixed effect estimator with transformations

# 2b - Show order of operation is important and explain why

# 2b - Change with imbalance?

# 3 - Regression analysis

Import Data
```{r}
#setwd("~/Dropbox/Berkeley_tings/Fall 2018/ARE213/Problem Sets/SharedFiles/are213/PS1b")
#setwd("C:\\Users\\will-\\Desktop\\are213\\PS1b")
setwd("C:\\Users\\Will\\Desktop\\are213\\PS1b")

dat <- read.dta("ps1.dta")


```




# 3a - Pooled bivariate OLS



```{r 3a}


```

# 3b - Standard errors for bivariate OLS



```{r 3b}


```


# 3c - between estimator



```{r 3c}


```

# 3d - random effects



```{r 3d}


```

# 3e - standard errors from RE



```{r 33}


```

# 3f - FE estimator



```{r 3f}


```

# 3g - stability of FE



```{r 3g}


```




